% mainfile: ../praca_magisterska_orbifoldy.tex
\chapter{Reduction to arithmetical questions}\label{reduction_to_arithmetical}
Reductions presented in this chapter will be more in the spirit of chapter \ref{order structure}, 
in the sense that 
%We will see (in the observation \ref{boils_down}) that the problem of determining this boils
% down to the 
% analysis of all 
%the possible 
%values of the expressions:
%\begin{equation}
%2 - \sum_{i=1}^n \frac{I_i-1}{I_i}
%\end{equation}
%and 
%\begin{equation}
%1 - \sum_{j=1}^m \frac{b_j-1}{2b_j},
%\end{equation}
%where $I_i, b_j$ varies over $\mathbb{N}_{>0} \cup \{\infty\}$. \\
%As
%\begin{equation}
%2 - \sum_{i=1}^n \frac{I_i-1}{I_i} = 2 - n + \sum_{i=1}^n \frac{1}{I_i}
%\end{equation}
%and 
%\begin{equation}
%1 - \sum_{j=1}^m \frac{b_j-1}{2b_j} = 1 - m + \sum_{j=1}^m \frac{1}{2b_j},
%\end{equation}
%some questions about the spectrum are equivalent to some regarding Egyptian fractions. 
%More on this connection is discussed in \ref{Egyptian_fractions}.
%\\[4pt]
%\textbf{Disclaimer}\\
for now, until chapter \ref{counting occurrences} 
%named "Counting occurrences" 
, we will not pay attention 
to how many orbifolds have the same Euler orbicharacteristic, only whether a particular 
number is an \Eoc\ for at least one orbifold or not. 

In chapter \ref{counting occurrences} we will explain how these reductions will be relevant 
to the discussion held there. 
%Let us note, that 
%Euler orbicharacteristic does not depend on the cyclic order of points on 
%the components of the boundary. 
%Because of that and since Euler orbicharacteristic does not depend on the cyclic order 
%of points on the components of the boundary we introduce an extension of a notation from 
%\cite{Conway2008}. 

%%We will write $\ast \{a,b,c,d,\cdots\}$ to denote not a particular orbifold, but a type 
%%of orbifold that have a dihedral points on a component of the boundry of orders 
%$a,b,c,d,\cdots$, but 
%%in any order. \\ 

%We will write $\ast \{a,b,c,d,\cdots\}$ to denote a type of a boundary (of an orbifold) that have 
%kaleidoscopic points of periods $a,b,c,d,\cdots$, but in any order. \\


%From what we wrote above (that Euler orbicharacteristic does not depend on the cyclic order 
%of points on the components of the boundary), we can see that Euler 
%orbicharacteristic is well defined 
%%on such types of orbifolds. \\ 
%when we specify only such a type of the components of the boundary of an orbifold and not 
%a particular cyclic order.  \\

\section{Reductions of cases}
The aim of following reductions is to make it easier to answer the question of which 
points lie in $\spe$ and which not. 

The first aspect of the structure of $\spe$ that we would like to simplify is that it is 
the sum of $\spebr{M}$, for 
every two dimensional manifold $M$. 
\begin{equation}
\spe = \bigcup_{M\textrm{ : 2d manifold}} \spebr{M}.
\end{equation}

%In this section we want to make some reductions to 
%limit number of cases that we will be dealing with. 

We aim to find a minimal set $\mathcal{M}$ of base manifolds 
%that "covers all the cases" i.e.  
such that:
\begin{equation}\label{minimal calM}
\spe = \bigcup_{M\in\mathcal{M}}\spebr{M}.
\end{equation}

% any for any $x \in \spe$ there 
%is an orbifold $O$ with a base manifold from $B$ such that $\cho{O} = x$.
%We aim to limit the number of base manifolds as much as possible while keeping entire spectrum. 
It will turn out that $\mathcal{M} = \{S^2, D^2\}$ satisfies \ref{minimal calM} and 
that both $S^2$ and $D^2$ are necessarily. 
%there are
% it can be done such that we are left with $B = \{S^2, D^2\}$ and there is 
%no futher reductions possible. 

%For this chapter we will consider orbifolds according to a definition from 
%(\ref{Disk_and_sphere_with_defects}). \\

\section{Sufficiency of $S^2$ and $D^2$}\label{sufficiency of D2 and S2}

Given an orbifold $O_1$, we want to perform some operations from \ref{Operations} on it, 
such that the resulting orbifold 
$O_2$ will have the same Euler orbicharacteristic, but the base manifold of $O_2$ would 
be $S^2$ or $D^2$. We would then say, that $O_1$ got reduced to $O_2$.
In following subsection, we allow only such operations, that do not 
change \Eoc. When writing that we "can" do something we mean that there is 
possible one of the operations from \ref{Operations}.  %have as 
%big Euler characteristic as possible. 

The Euler characteristic of base manifold depends only on the number of handles, cross caps 
and boundary components. And, as stated in \ref{E_orb} it is: 
\begin{equation}
2-2h-c-b,
\end{equation}
%\smalltodoII{opisać i wybrać oznaczenia}
for $h$ - number of handles, $c$ - number of cross-caps, $b$ - number of boundary components. 

For every such a manifold feature we want to find an orbifold features with the same 
Euler orbicharacteristic delta. 

We will take two approaches, depending on whether the orbifold in question has a boundary or not.

%To do this we want to eliminate handles 
%Let us observe, that:
\subsection{Orbifold without boundary}
We can observe that:
\begin{align}
\hspace{3cm}\Delta(\circ) =& \hspace{-1cm} &-2& \hspace{-1cm} &= \Delta(2^4) \\
\hspace{3cm}\Delta(\times) =& \hspace{-1cm} &-1& \hspace{-1cm} &= \Delta(2^2)
\end{align}
From this we can see that we can remove handles and cross-caps from any orbifold without 
the boundary. 
After such reductions we are left with a $S^2$ orbifold with all orbipoints being rotational 
in the interior.

\subsection{Orbifold with boundary}\label{orbifold_with_boundary}
We can observe that:
\begin{align}
\hspace{3cm}\Delta(\circ) =& \hspace{-1cm} &-2& \hspace{-1cm} &= \Delta((^\ast 2)^8) \\
\hspace{3cm}\Delta(\ast) =& \hspace{-1cm} &-1& \hspace{-1cm} &= \Delta((^\ast 2)^4) \\
\hspace{3cm}\Delta(\times) =& \hspace{-1cm} &-1& \hspace{-1cm} &= \Delta((^\ast 2)^4)
\end{align}
From this we can see that we can remove handles and cross-caps 
from any orbifold with a boundary. 
We can also remove all boundary components except one.         
We can further observe that:
\begin{align}
\hspace{3cm}\Delta(n) =& \hspace{-1cm} &\frac{n-1}{n}& \hspace{-1cm} 
&= 2\frac{n-1}{2n} &= \Delta((^\ast n)^2)\label{third_reduction}
\end{align}
From this we see that we can remove all the rotational orbipoints in favor for 
dihedral orbipoints.
After such reductions we are left with a $D^2$ orbifold with all orbipoints being dihedral on 
the boundary or being reflectional on the boundary. 
%\smalltodoII{reflectional?}

As a fact not necessary for our reductions, but interesting on its own, we can furthermore, 
observe that:
\begin{observation}
If $O_1$ has not $S^2$ as its base manifold it can be reduced to a $D^2$-orbifold.
\end{observation}
\subsubsection{Proof.}
If $O_1$ has not $S^2$ as its base manifold $M$, then $M$ has at least one handle or a cross-cup. We can observe that:
\begin{align}
\hspace{3cm}\Delta(\circ) =& \hspace{-1cm} &-2& \hspace{-1cm} &= \Delta(\ast (^*2)^4) \\
\hspace{3cm}\Delta(\times) =& \hspace{-1cm} &-1& \hspace{-1cm} &= \Delta(\ast).
\end{align}
From this we have that the handle or the cross-cap can be replaced by a boundary component  
and some number of boundary orbipoints. After this reduction, we can proceed with all the 
other reductions from the \ref{orbifold_with_boundary} and obtain an $D^2$-orbifold 
with the same \Eoc as the original one. $_\square$

%\begin{align}
%\Delta(n) =& &\frac{n-1}{n}&  
%&= \Delta((^\ast n)^2)\label{third_reduction}
%\end{align}

%From this we can conclude that every Euler orbicharacteristic can be obtained 
%by an orbifold with base manifold $S^2$ or $D^2$. 
%Examples of rational numbers from $\speS \setminus \speD$ and $\speD \setminus \speS$ are:
%We will provide examples 
%
%\ref{counting_occurences}

%From this we can conclude, that every Euler orbicharacteristic can be obtained 
%by an orbifold of signature of a type ($n$ and $m$ are arbitrary):

%\begin{align*}
%I_1I_2\cdots I_n & \textrm{\ or} \\
%\ast b_1b_2\cdots b_m &.
%\end{align*}

%Let us denote the set of all possible Euler orbicharacteristics of orbifolds of the form 
%$I_1I_2\cdots I_n$ by $\speS$ and the set 
%of all possible Euler orbicharacteristics of orbifolds of the form $\ast b_1b_2\cdots b_m$ 
%as $\speD$. 
%So we have that $\spe = \speD \cup \speS$. \\
%% Let us denote the set of all possible \Eoc s of two-dimentional orbifolds as $\spe$. \\
%Let us also observe that the order type and topology of $\speS$ and $\speD$ are 
%the same since 

%\begin{equation}\label{times_two_fact}
%2\speD=\speS
%\end{equation}
%and multiplying by $2$ is the order preserving homeomorphism of $\mathbb{R}$. \\[16pt]

%Now we can make aforementioned observation:
%\todo{dopisać to czytelniej}
%w szeczgólności żeby było jasne jaką redukcję robimy

The results of our reductions, can be summarised as:
%In the terms of set relations:
\begin{observation}\label{sum of spectras}
If two-dimensional manifold $M$ has no boundary, then
%\begin{equation} 
%\spe{M} = \chi(M) - (\speS - 2) 
%\end{equation}
%and
\begin{equation} 
\spebr{M} \subseteq \speS 
\end{equation} 

If, in addition, $M \neq S^2$, then 
\begin{equation} 
\spebr{M} \subseteq \speD. 
\end{equation}

\end{observation}
%\textbf{Proof} \\
%Let $M$ be a two-dimentional manifold with no boundry, from \ref{\Eoc_as_a_sum} we have, that: 
%\begin{equation}
%\spe{M} = \{\chi(M) - \sum_{i=1}^n 
%\frac{I_i-1}{I_i}\ |\ n \in \mathbb{N} \land \forall_i I_i \in \mathbb{N} \cup \infty\}.
%\end{equation}
%And from 
%$\chi(M) - (\speS - 2)$
\begin{observation}
If two-dimensional manifold  $M$ has a boundary, then 
%\begin{equation}
%\spe{M} = \chi(M) + (\speD - 1)
%\end{equation}
%and 
\begin{equation}
\spebr{M} \subseteq \speD
\end{equation}
\end{observation}
%\textbf{Proof} \\
%\begin{equation}
%\{\chi(M) - \sum_{j=1}^m 
%\frac{b_j-1}{2b_j}\ |\ m \in \mathbb{N} \land \forall_j b_j \in \mathbb{N} \cup \infty\} 
%\end{equation}
%\begin{observation}
%If two-dimentional manifold  $M$ has no boundry, then $\spe{M} \subseteq \speS$. 
%If, in addition, $M \neq S^2$, then 
%$\spe{M} \subseteq \speD$.
%\end{observation}
%\begin{observation}
%If two-dimentional manifold  $M$ has a boundry, then $\spe{M} \subseteq \speD$.
%\end{observation}

%In the terms of arithmetical expressions:
%\todo{sums are the form od sped and spes}

\begin{corollary}
We have that $\spe = \speS \cup \speD$.
\end{corollary}

\begin{observation}\label{only dihedral} 
If a two-dimensional manifold $M$ has a boundary, then:
\begin{equation}
\spebr{M} = \spe^{d}(M).
\end{equation}
\end{observation}

We will postpone our discussion of necessity of both $S^2$ and $D^2$ to 
\ref{neccessity of d2 and s2}, after 
the 
section \ref{Reduction to arithmetical questions section} which will provide us 
with more convenient language. 

\section{Reduction to arithmetical questions}\label{Reduction to arithmetical questions section}
As written in \ref{spectra}, we can express an \Eoc of a $M$-orbifold $O$ as:
\begin{equation}
\cho{O} = \chi (M) - \sum_{i=1}^n \frac{r_i-1}{r_i} - \sum_{j=1}^m \frac{d_j-1}{2d_j},
\end{equation}
where $r_i$ and $d_j$ are degrees of the, respectively, rotational and dihedral orbipoints 
of $O$.

From this we can express $\spebr{M}$ as:
\begin{align}
\spe(M) = \chi (M) &-\left\{\sum\limits_{i=1}^n \frac{r_i-1}{r_i}\ \big|\ n\in\mathbb{N}_0,\ 
r_i\in\mathbb{N}_{>0}\cup \{\infty\}\right\}+ \\
&- \left\{\sum\limits_{j=1}^m\frac{d_j-1}{2d_j}\ 
\big|\ m\in\mathbb{N}_0,\ d_j\in\mathbb{N}_{>0}\cup \{\infty\}\right\} .
\end{align}

As from \ref{sufficiency of D2 and S2} we know that $\spe = \speS \cup \speD$, and that 
$\chi (S^2) = 2$ and $\chi (D^2) = 1$, we can express $\spe$ as a sum ($\cup$) of two sets:
\begin{equation}
2 -\left\{\sum\limits_{i=1}^n \frac{r_i-1}{r_i}\ \big|\ n\in\mathbb{N}_0,\ 
r_i\in\mathbb{N}_{>0}\cup \{\infty\}\right\} = \spe(S^2)
\end{equation}
and
\begin{equation}
1 - \left\{\sum\limits_{j=1}^m\frac{d_j-1}{2d_j}\ 
\big|\ m\in\mathbb{N}_0,\ d_j\in\mathbb{N}_{>0}\cup \{\infty\}\right\} = \spe(D^2) .
\end{equation}

From this we see, that the core of understanding $\spe$ through arithmetical viewpoint 
is to understand possible values of expression:
%We now know, that $\spe = \speS \cup \speD$. To determin 

%\begin{observation}\label{boils_down}
%From above reductions we can conclued that our problem boiles down to the analysis of all
% the possible 
%values of the expressions:
\begin{equation}\label{S2_sum}
2 - \sum_{i=1}^n \frac{r_i-1}{r_i}
\end{equation}
and 
\begin{equation}\label{D2_sum}
1 - \sum_{j=1}^m \frac{d_j-1}{2d_j},
\end{equation}
with $r_i$ and $d_j$ ranging over $\mathbb{N}_{>0}\cup \{\infty\}$.
%, with a convention 
%that $\varphi(\infty) \coloneqq \lim\limits_{k\to \infty} \varphi(k)$. 
%\end{observation}

As $\Delta(\infty) = 1 = \Delta(2^2)$ and $\Delta(^*\infty) = \frac{1}{2} = \Delta((^*2)^2)$, 
we could perform further reductions to have an orbifold with 
particular orbicharacteristic without cusps (if needed) and then (after these reductions) 
we can analyse only expressions with $r_i$ and $d_j$ ranging over $\mathbb{N}_{>0}$ and 
they will still give us full spectrum. 
However, as stated later, it will be more convenient to us to include orbifolds with cusps 
so we are stating this observation only as a side remark.
%for readers information. 
%The fact that this agrees with the definition of the \Eoc\ on the geometrical terms was 
%addressed in \ref{extended_Euler_orbicharacteristic}. 

\section{Hurwitz theorem}\label{największy orbifold}
%\subsection{Hurwitz theorem}
One of the well known facts about two-dimensional orbifolds comes from \cite{Hurwitz1893} 
(seite 424).
Although formulated in different language, it states, that the orbifolds 
with hyperbolic structure can have an \Eoc\ at most $-\frac{1}{84}$. 
Here we will present the proof of this result with the language used in this thesis.
%We will now prove this result 
%in a straighforward way. 
%\todo{dopisać}
\begin{theorem}
If a two dimensional orbifold admits hyperbolic structure, then the maximal \Eoc\ it can 
have is $-\frac{1}{84}$ and the only orbifold that realises this \Eoc\ is $*2\ 3\ 7$.
\end{theorem}
\subsubsection{Proof.}
From \ref{sufficiency of D2 and S2} we know, that to check whether $-\frac{1}{84}$ is maximal 
negative \Eoc\ a two dimensional orbifold can have, we only need to check possible 
\Eoc s of $S^2$ orbifolds and $D^2$ orbifolds. 

Firstly, we will show, that $*2\ 3\ 7$ has biggest negative \Eoc\ from all $D^2$ orbifolds 
and its the only one with this \Eoc\ from $D^2$ orbifolds.

We have that 
\begin{equation}
\cho{*2\ 3\ 7} = 1 - \frac{1}{4} - \frac{2}{6} - \frac{6}{14} = -\frac{1}{84}. 
\end{equation}
From \ref{spectra} we know, that $\speD = \spe^d(D^2)$, so we can consider only dihedral 
orbipoints. 
Let us observe, that since 
$\max\{\Delta(^*n)\ |\ n \in \mathbb{N}_{>0} \cup \{\infty\}\} = \frac{1}{2}$, we have, that 
to have \Eoc\ $<0$, 
$D^2$ orbifold has to have at least $3$ orbipoints.
%Let us observe
 
Let us observe, that for any $D^2$ orbifold that have $5$ orbipoints or more, it has an \Eoc 
equal at most $1 - 5\frac{1}{4} = -\frac{1}{4} < -\frac{1}{84}$. So we can restrict our search 
only to orbifolds with at most $4$ orbipoints. 

Let us observe, that $*2\ 2\ 2\ 3$ and $* 3\ 3\ 4$ are $D^2$ orbifolds with the biggest 
negative \Eoc\ among orbifolds, respectively, with four orbipoints, and, without any 
point of degree $2$. The proof of this observation is, that any other orbifold of these 
respective kinds, would need to have all degrees of orbipoints (when ordered in increasing manner) 
pairwise $\geq$ than $*2\ 2\ 2\ 3$ or $* 3\ 3\ 4$.
%Let us observe, that id $D^2$ orbifold have at least one rotational orbipoint, then 
%the smallest possible \Eoc\ is 
%\begin{equation}
%\cho{2 * 2 3} = 1 - \ 
%\end{equation}

Let us observe, that 
%the orbifold wsith the greatest negative \Eoc, that 
%have no orbipoints of degree $2$, that is, $* 3\ 3\ 4$, have \Eoc\ 
\begin{equation}
\cho{*2\ 2\ 2\ 3} = 1 - \frac{1}{4} - \frac{1}{4} - \frac{1}{4} - \frac{2}{6} = -\frac{1}{12} 
< -\frac{1}{84}
\end{equation}
and
\begin{equation}
\cho{*3\ 3\ 4} = 1 - \frac{2}{6} - \frac{2}{6} - \frac{3}{8} = -\frac{1}{24} < -\frac{1}{84},
\end{equation}
From this, we can conclude, that we can restrict our search only to orbifolds with 
exactly $3$ orbipoints, where at least one of them is equal to $2$. 

Let us observe, that for such orbifold to have negative \Eoc, it needs to have at two orbipoints 
of order at least $3$, otherwise it has \Eoc\ at most 
\begin{equation}
\cho{*2\ 2\ \infty} = 1- \frac{1}{4} - \frac{1}{4} - \frac{1}{2} = 0. 
\end{equation}

Further, let us observe, that $*2\ 4\ 5$ has the greatest negative \Eoc\ among orbifolds 
that have no orbipoint of degree $3$. The proof of this observation is similar to the proof 
of the previous one -- any other orbifold of this kind, 
would need to have all degrees of orbipoints (when ordered in increasing manner) 
pairwise $\geq$ than $*2\ 4\ 5$.

Let us observe, that
%We have that 
\begin{equation}
\cho{*2\ 4\ 5} = 1 - \frac{1}{4} - \frac{3}{8} - \frac{4}{10} = -\frac{1}{40} < - \frac{1}{84}. 
\end{equation}
From this, we conclude, that we can restrict our search to the orbifolds of 
the form $*2\ 3\ n$. 
We can also observe, that all orbifolds of the form $*2\ 3\ n$ have unique \Eoc\ 
among this group.
%and it has to be unique. 
The one with the biggest \Eoc\ among them is $*2\ 3\ 7$. 

Let us observe that no $D^2$ orbifold with rotational orbipoints can't have such \Eoc. 
For the sake of contradiction, let us assume that there is some orbifold 
$r_1 \cdots r_n * d_1 \cdots d_m$, with $n\neq 0$, with \Eoc\ equal to $-\frac{1}{84}$. 
However, then also $*r_1r_1 \cdots r_nr_nd_1\cdots d_m$, 
with only dihedral orbipoints, 
would have \Eoc\ equal to $-\frac{1}{84}$. However, $*2\ 3\ 7$ is unique one with this \Eoc\ 
and it have no repeated degree, so it can't be expressed in the form 
$*r_1r_1 \cdots r_nr_nd_1\cdots d_m$ 
with $n\neq 0$. 

Let us also observe, that the same argument shows that $\mathbb{R}P^1$ orbifolds 
can't have \Eoc\ equal to $-\frac{1}{84}$, since $\chi{\mathbb{R}P^1} = \chi{D^2}$ 
and $\mathbb{R}P^1$ has no boundary, so $\mathbb{R}P^1$ orbifold can have only rotational 
orbipoints. 
 
Now, we will prove, that no $S^2$ orbifold has \Eoc\ $-\frac{1}{84}$. 
%From this, 
Since we have \ref{2times homeomorphism} 
%it will 
%it folows, that no $S^2$ orbifold 
%has greater negative \Eoc. 
%The argument goes as folows:
We can perform following argument:  
 
For the sake of contradiction let us assume, that 
$-\frac{1}{84} \in \speS$, then, from \ref{2times homeomorphism} we know, that 
$\frac{1}{2}\left(-\frac{1}{84}\right)\in \speD$. This is a contradiction as 
$0 > \frac{1}{2}\left(-\frac{1}{84}\right) > -\frac{1}{84}$. 
As such, we ruled out all manifolds with Euler characteristic $>0$. 

For 
%What is left to prove, is that $-\frac{1}{84}$ is not an \Eoc\ for an $M$ orbifold 
%for any 
two dimensional manifolds 
%$M$ 
%We can rule out all manifolds 
with Euler characteristic $\leq 0$, we have that orbifolds having them as a base manifolds 
have \Eoc at most $-\frac{1}{4} < -\frac{1}{84}$. 
%From the arguments above, we can also conclude, that we can also rule out 
%$\mathbb{R}\mathrm{P}^1$ -- the only 
%other manifold left under our consideration, as orbifolds corresponding to it can only have 
%rotational orbipoints. 
$_\square$ 

\section{Necessarily of $S^2$ and $D^2$}\label{neccessity of d2 and s2}
As we know from \ref{Operations} adding an orbipoint to a manifold decreases it's 
orbicharacteristic. As $S^2$ has the highest Euler characteristic: $2$ of all 
two dimensional manifolds, there is no other orbifold with \Eoc\ equal to $2$. 
$S^2$ is then necessary to include $2$. 

As known from \ref{największy orbifold}, the number $-\frac{1}{84}\in\speD$, it is 
the greatest negative 
\Eoc any two dimensional orbifold can have and $-\frac{1}{84}\not\in\speS$. 
%We will now show, that 
%$-\frac{1}{84} \not\in \speS$. For the sake of contradiction let us assume, that 
%$-\frac{1}{84} \in \speS$, then, from \ref{2times homeomorphism} we know, that 
%$\frac{1}{2}\left(-\frac{1}{84}\right)\in \speD$. This is a contradiction as 
%$0 > \frac{1}{2}\left(-\frac{1}{84}\right) > -\frac{1}{84}$. 
$_\square$  

Further examination of connections between $\speD$ and $\speS$ is performed in \ref{D_and_S}.






